Wikipedia Text Mining and Surprisal Analysis

### 2019-06-09 surprisal/MacWhinney poster/poster_plotting.Rmd
unigram letter surprisals

---
title: Positional surprisal (1-3grams) in CHILDES corpora
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

@@

label_data <- tibble(language = "English",
                     speaker_role = unique(data$speaker_role),
                     word_order = c(2, 3, 4),
                     s = c(6, 5, 4),
                     context = "Unigram") %>%
  mutate(context = factor(context, levels = c("Unigram", "Trigram")),
         language = factor(language, 
                           levels = c("English", "Spanish", 
                                      "Mandarin", "Japanese")))

pdf("surprisals.pdf", width = 10, height = 5.5)
data %>%
  filter(length == 8) %>%
   ggplot(aes(x = word_order, y = s,
                        color = speaker_role,
                        label = speaker_role)) +
  facet_grid(context ~ language, scales = "free") +
  geom_point(size = 1.5) +
  geom_line(size = 1.5) +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.position = "none") +
  geom_text(data = label_data, size = 6) + 
  labs(x = "Utterance Position", y = "Surprisal")
dev.off()
```

Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.

```{r hard-coded language unigrams}
child_utterances <- prov_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_unigrams <- child_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

child_surprisals <- child_utterances %>%
  left_join(child_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

adult_surprisals <- adult_utterances %>%
  left_join(adult_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

```

```{r new bigrams}
adult_bigrams_prep <- adult_utterances %>% 
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x / n.y) %>%
  select(lag_word, word, cond_p)

adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  left_join(adult_unigrams) %>%
  left_join(adult_bigrams_prep) %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(mean_s = mean(s)) 
```

```{r bigrams}
adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x/n.y)

bigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(adult_bigrams) %>%
  left_join(adult_unigrams)

bigram_surprisals <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

```

```{r trigrams}
adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n / n.x) %>%
  select(-n, -n.x, -n.y, -p, -cond_p)

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>% 
  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(adult_unigrams)

trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)
```

>>>>

---
title: Plotting surprisal in CHILDES corpora and Switchboard English corpus
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

@@

label_data <- tibble(language = c("Spanish", "Spanish", "English"),
                     speaker_role = unique(data$speaker_role),
                     word_order = c(2, 4, 2),
                     s = c(6.5, 5, 6.5),
                     context = "Unigram") %>%
  mutate(context = factor(context, levels = c("Unigram", "Trigram")),
         language = factor(language, 
                           levels = c("English", "Spanish", 
                                      "Mandarin", "Japanese")))

pdf("surprisals.pdf", width = 10, height = 5.5)
data %>%
  filter(length == 8) %>%
   ggplot(aes(x = word_order, y = s,
                        color = speaker_role,
                        label = speaker_role)) +
  facet_grid(context ~ language, scales = "free") +
  geom_point(size = 1.5) +
  geom_line(size = 1.5) +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.position = "none") +
  geom_text(data = label_data, size = 6) + 
  labs(x = "Word Position", y = "Average Information")
dev.off()
```

### 2019-06-09 surprisal/wiki_surprisal.Rmd
unigram letter surprisals

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(missMDA)
library(cluster)
library(widyr)
library(lsa)
library(magrittr)
library(broom)
library(here)
library(feather)
library(FactoMineR)
library(feather)
library(janitor)

@@

```{r make dendrogram}
wiki <- read.csv(here("Data/relative_unigrams.csv")) %>%
  select(-X) %>% 
  column_to_rownames("Language")

wiki %>% 
  dist() %>% 
  hclust() %>% 
  plot(hang = -1, 
       main = "Dendrogram of slope embeddings")

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lingtypology)
library(knitr)
library(missMDA)
library(cluster)
library(widyr)
library(lsa)
library(magrittr)
library(broom)
library(here)
library(feather)
library(FactoMineR)
library(feather)
library(janitor)
library(ggdendro)

@@

```{r make dendrogram}
wiki <- read.csv(here("Data/wikipedia_0context.csv")) %>%
  select(-X) %>% 
  column_to_rownames("Language")

wiki %>% 
  dist() %>% 
  hclust() %>% 
  ggdendrogram(rotate = FALSE, size = .1) + 
  ylim(0, .10)

  plot(hang = -1, 
       main = "Dendrogram of slope embeddings")
